Reproducible Research Peer Assessment 01
========================================================

Load the plyr library

```{r}
library(plyr)
library(dplyr)
library(lattice)
```


Loading and pre-processing the data
-

1 Load the data

```{r "read in the data"}
#download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "activity.zip", mode="wb")
#unzip("activity.zip")
rawData <- read.csv("activity.csv", header=TRUE)
```

2 Process/transform the data (if necessary) into a format suitable for your analysis

Not required in my opinion



What is mean total number of steps taken per day?


1 Calculate the total number of steps taken per day

```{r "Sum of steps per day"}
sumOfSteps <- ddply(rawData,~date, summarise, sum=sum(steps))
```
Load the dplyr library for data manipulations

```{r, warning=TRUE, message=FALSE}
library("dplyr")
sumOfDailySteps <- summarise(group_by(rawData, date), sum=sum(steps))
```


2 Make a histogram of the total number of steps taken each day

```{r "Plot of steps taken per day"}
hist(sumOfDailySteps$sum, 
     breaks=20,
     main="Histogram of steps taken each day",
     xlab="Number of steps",
     ylim=c(0,20), 
     xlim=c(0,25000))
```

3 Calculate and report the mean and median of the total number of steps taken per day


```{r}
meanOfSteps <- mean(sumOfDailySteps$sum, na.rm=T)

medianOfSteps <- median(sumOfDailySteps$sum, na.rm=T)
```

**Mean of steps:** `r I(format(meanOfSteps, digits=7))`

**Median of steps:** `r I(format(medianOfSteps, digits=7))`


What is the average daily activity pattern?
-

1 Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)


```{r "Mean of Steps without NA"}
meanOfStepsNoNA <- ddply(na.omit(rawData),~interval, summarise, steps=mean(steps))
plot(meanOfStepsNoNA$interval, meanOfStepsNoNA$steps, type="l")
```

2 Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
tail(meanOfStepsNoNA[ order(meanOfStepsNoNA[,2]),], n=1)
```

Imputing missing values
-

1 Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
summary(rawData$steps==NA)
```

2 Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r "Plot with NA replacement"}
medianOfStepsNoNA <- ddply(na.omit(rawData),~interval, summarise, steps=median(steps))
plot(meanOfStepsNoNA$steps, type="l")
lines(medianOfStepsNoNA$steps)
```

My strategy would be to replace with the mean value of that timepoint

3 Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
cleanData <- rawData

for (ii in 1:dim(rawData)[1]){ 
  if (is.na(rawData$steps[ii])){
    int <- rawData$interval[ii]
    cleanData$steps[ii] <- round(meanOfStepsNoNA[meanOfStepsNoNA$interval==int,]$steps)
  }
}
```

4 Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from  the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r "Histogram of Steps per Day from cleaned data"}
sumOfCleanedSteps <- ddply(cleanData,~date, summarise, sum=sum(steps))

hist(sumOfCleanedSteps$sum, 
     breaks=20, 
     main="Histogram of cleaned steps taken each day",
     xlab="Number of steps",
     ylim=c(0,20), 
     xlim=c(0,25000))

meanOfCleanedSteps <- ddply(cleanData,~date, summarise, steps=mean(steps))
ddply(cleanData,~date, summarise, steps=mean(steps))
medianOfCleanedSteps <- ddply(cleanData,~date, summarise, steps=median(steps))
ddply(cleanData,~date, summarise, steps=median(steps))
```

Are there differences in activity patterns between weekdays and weekends?
Yes, the graphs show a slight difference in the pattern of steps.
-

```{r "Plot comparing Weekend & Weekday step patterns"}
cleanData$DoW <- ifelse( weekdays(as.Date(cleanData$date)) == "Saturday"| weekdays(as.Date(cleanData$date)) =="Sunday" , "weekend" , "weekday" )

tmp <- ddply(cleanData,~interval*DoW, summarise, steps=mean(steps))

xyplot(tmp$steps~tmp$interval | as.factor(tmp$DoW), type="l", layout=c(1,2))
```



